---
title: "Comparison of ANCOVA.xx with Base R Functions"
output: rmarkdown::html_vignette
vignette: >
  %\\VignetteIndexEntry{Comparison of ANCOVA.xx with Base R Functions}
  %\\VignetteEngine{knitr::rmarkdown}
  %\\VignetteEncoding{UTF-8}
---

# Comparison of ANCOVA.xx with Base R Functions

This vignette compares the functionality, efficiency, and flexibility of the ANCOVA.xx package with base R functions such as lm and anova. Examples use both real-world and simulated datasets.

```{r}
# Load required packages
library(ANCOVA.xx)
if (!requireNamespace("bench", quietly = TRUE)) {
  install.packages("bench")
}
library(bench)
# Load the dataset
data(mtcars)

```

## Real-World Dataset Comparison: mtcars

### Regression: lm vs. run_regression

```{r example, echo = TRUE}
# Using updated run_regression
custom_model <- run_regression(mtcars, response = "mpg", predictors = c("wt", "hp"))

# Using lm
base_model <- lm(mpg ~ wt + hp, data = mtcars)

# Compare coefficients
print(all.equal(as.numeric(custom_model$coefficients), as.numeric(coef(base_model)), tolerance = 1e-5))

# Compare R-squared
print(all.equal(custom_model$r_squared, summary(base_model)$r.squared, tolerance = 1e-5))

bench_coefficients <- bench::mark(
  run_regression = {
    custom_model <- run_regression(mtcars, response = "mpg", predictors = c("wt", "hp"))
    as.numeric(custom_model$coefficients)  # Extract coefficients
  },
  lm = {
    base_model <- lm(mpg ~ wt + hp, data = mtcars)
    as.numeric(coef(base_model))  # Extract coefficients
  },
  iterations = 100
)

print("Coefficient Computation Benchmark:")
print(bench_coefficients)
```

### ANOVA: anova vs. compare_anova
```{r}
# Fit models for ANOVA
base_model1 <- lm(mpg ~ wt, data = mtcars)
base_model2 <- lm(mpg ~ wt + hp, data = mtcars)

# Base R ANOVA
base_anova <- anova(base_model1, base_model2)
print("Base R ANOVA Results:")
print(base_anova)

# Using compare_anova
custom_model1 <- run_regression(mtcars, response = "mpg", predictors = c("wt"))
custom_model2 <- run_regression(mtcars, response = "mpg", predictors = c("wt", "hp"))
custom_anova <- compare_anova(models = list(custom_model1, custom_model2))
print("Custom ANOVA Results:")
print(custom_anova)

# Compare F-values
f_value_comparison <- all.equal(
  custom_anova$F_value[2],
  base_anova$F[2],
  tolerance = 1e-5
)
print(paste("F-value comparison:", f_value_comparison))

# Compare p-values
p_value_comparison <- all.equal(
  custom_anova$p_value[2],
  base_anova$`Pr(>F)`[2],
  tolerance = 1e-5
)
print(paste("P-value comparison:", p_value_comparison))


# Benchmark ANOVA (numeric values only)
bench_anova <- bench::mark(
  compare_anova = {
    compare_anova(models = list(custom_model1, custom_model2))$F_value[2]
  },
  base_anova = {
    anova(base_model1, base_model2)$F[2]
  },
  iterations = 100
)


print("ANOVA Benchmark:")
print(bench_anova)
```

### ANCOVA: aov vs. compare_ancova

```{r}


custom_ancova <- compare_ancova(
  models = list(
    run_regression(mtcars, response = "mpg", predictors = c("wt")),
    run_regression(mtcars, response = "mpg", predictors = c("wt", "hp"))
  ),
  covariates = c("wt", "hp"),
  data = mtcars
)
print(custom_ancova)
base_ancova <- aov(mpg ~ wt + hp, data = mtcars)
summary_base_ancova <- summary(base_ancova)

# Compare F-values and p-values for "hp"
base_f_value_hp <- summary_base_ancova[[1]]$`F value`[2]
base_p_value_hp <- summary_base_ancova[[1]]$`Pr(>F)`[2]

custom_f_value_hp <- custom_ancova$F_value[custom_ancova$Covariate == "hp" & custom_ancova$Model == "Model_2"]
custom_p_value_hp <- custom_ancova$p_value[custom_ancova$Covariate == "hp" & custom_ancova$Model == "Model_2"]

print(paste("Base F-value (hp):", base_f_value_hp))
print(paste("Custom F-value (hp):", custom_f_value_hp))

print(paste("Base p-value (hp):", base_p_value_hp))
print(paste("Custom p-value (hp):", custom_p_value_hp))


# Base R ANCOVA
base_ancova <- aov(mpg ~ wt + hp, data = mtcars)
summary_base_ancova <- summary(base_ancova)

# Using compare_ancova
custom_ancova <- compare_ancova(
  models = list(custom_model1, custom_model2),
  covariates = c("wt", "hp"),
  data = mtcars
)


# Extract numeric values for hp
base_f_value_hp <- summary_base_ancova[[1]]$`F value`[2]
base_p_value_hp <- summary_base_ancova[[1]]$`Pr(>F)`[2]

custom_f_value_hp <- custom_ancova$F_value[custom_ancova$Covariate == "hp" & custom_ancova$Model == "Model_2"]
custom_p_value_hp <- custom_ancova$p_value[custom_ancova$Covariate == "hp" & custom_ancova$Model == "Model_2"]

# Compare numeric values
f_value_hp_comparison <- all.equal(custom_f_value_hp, base_f_value_hp, tolerance = 1e-5)
p_value_hp_comparison <- all.equal(custom_p_value_hp, base_p_value_hp, tolerance = 1e-5)

print(paste("F-value comparison for hp:", f_value_hp_comparison))
print(paste("P-value comparison for hp:", p_value_hp_comparison))

# Benchmark ANCOVA (numeric values only)
bench_ancova <- bench::mark(
  compare_ancova = {
    compare_ancova(models = list(custom_model1, custom_model2), covariates = c("wt", "hp"), data = mtcars)$F_value[custom_ancova$Covariate == "hp" & custom_ancova$Model == "Model_2"]
  },
  base_ancova = {
    summary(aov(mpg ~ wt + hp, data = mtcars))[[1]]$`F value`[2]
  },
  iterations = 100
)

print("ANCOVA Numeric Value Benchmark:")
print(bench_ancova)
```

### ANCOVA with Non-Nested Models

**Simulation**
```{r}
set.seed(456)
sim_data2 <- data.frame(
  y = rnorm(500, mean = 100, sd = 15),
  x1 = rnorm(500, mean = 50, sd = 10),
  x2 = rnorm(500, mean = 30, sd = 5),
  group = factor(sample(1:3, 500, replace = TRUE))
)
```
```{r}
# Base R ANCOVA
lm_model1 <- lm(y ~ x1, data = sim_data2)
lm_model2 <- lm(y ~ x1 + x2, data = sim_data2)
anova_results <- anova(lm_model1, lm_model2)
print(anova_results)


# Using ANCOVA.xx
model1_sim2 <- run_regression(sim_data2, response = "y", predictors = c("x1"))
model2_sim2 <- run_regression(sim_data2, response = "y", predictors = c("x1", "x2"))
ancova_results_sim2 <- compare_ancova(models = list(model1_sim2, model2_sim2), covariates = c("x1", "x2"), data = sim_data2)
print(ancova_results_sim2)
```
